---
title: "Exploding Boxplots"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(bpexploder)
```

## A Simple Widget

`bpexploder` represents my first foray into Html Widgets for R; it renders box-plots that explode upon mouse-click into jittered individual-value plots.  You have the option to configure tool-tips for the individual points.


## Installation and Usage

Install the package from Git Hub:

```{r eval = FALSE}
devtools::install_github("homerhanumat/bpexploder")
```



Call the function with hello-world defaults for the `iris` data:

```{r eval = FALSE}
bpexploder()
```

Click to explode boxes, double-click to restore.

##  Settings

`bpexploder` provides modest options for customization, as illustrated in the following example:

```{r eval = FALSE}
bpexploder(data = chickwts,
          settings = list(
            yVar = "weight",
            # default NULL makes one plot for yVar
            groupVar = "feed",
            levels = levels(with(chickwts,
                                 reorder(feed, weight, median))),
            # you could adjust the group lables ...
            levelLabels = NULL,
            # ... and the colors for each group:
            levelColors = NULL,
            yAxisLabel = "6-week weight (grams)",
            xAxisLabel = "type of feed",
            tipText = list(
              # as many os you like of:
              # variableName = "desired tool-tip label"
              # leave tipText at NULL for no tips
              weight = "weight")
            )
          )
```

## Sizing

By default the box-plot chart sizes itself as the offset-width of its grandparent node in the HTML DOM.  For some layout this might not be what you want, in which case you may direct `epexploder()` to track the offset-width of an existing DOM element.  For example, if paragraphs in your document have the desired width, then create an empty paragraph in your markdown like this:

```
<p id="reference"></p>
```

To set the widget-width that to a specific fraction of the element that contains the paragraph, you might try something like: 

```
<p id="reference" style="width: 70%"></p>
```

<p id="reference" style="width: 70%"></p>

Then call `bpexploder()` with the `referenceId` setting:

```{r eval = FALSE}
bpexploder(data = iris,
          settings = list(
            groupVar = "Species",
            levels = levels(iris$Species),
            yVar = "Petal.Length",
            tipText = list(
              Petal.Length = "Petal Length"
            ),
            referenceId = "reference"
            )
          )
```

## Known Issues

* Using the Leonids theme from the `prettydoc` R Markdown package, tooltips are not visible in Firefox or Safari.  (They do show in Chrome.)
* When `htmlwidgets::createWidget()` makes a widget it gives it a random Id number.  For some reason this can result in spurious warnings concerning the hidden varible `.Random.Seed`.  To work around this, set a seed in the `setup` chunk of your R Markdown document, e.g.:

    ```{r eval = FALSE}
    library(bpexploder)
    set.seed(5437)  # one way to avoid .Random.Seed warning from widgetId creation
    ```

## Working Examples

For a live version of this document see [this R Pub](http://rpubs.com/gccsc/bpexploder).

## Credits

The JavaScript library is based on Mathieu Caule's [D3 Exploding Boxplots](https://mcaule.github.io/d3_exploding_boxplot/), which I have modified slightly and updated for D3 Version 4. The tool-tips were created by [Justin Palmer](https://github.com/Caged), updated to D3v4 by [Dave Gotz](https://github.com/VACLab/d3-tip) and tweaked slightly by myself.
